Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
done for now, really

\author[mcgill]{Reza Adalat}
\author[mcgill]{Natacha Beck}
\author[mcgill]{\\Samir Das}
\author[isi]{Rafael Ferreira da Silva}
\author[mcgill]{Najmeh Khalili-Mahani}
\author[criugm]{Pierre-Olivier Quirion}
\author[mcgill]{Pierre Rioux}

\author[criugm]{Pierre Bellec}
\author[mcgill]{Alan C. Evans}

\address[mcgill]{McGill Centre for Integrative Neuroscience, Montreal Neurological Institute, McGill University, Canada.}
\address[creatis]{University of Lyon, CNRS, INSERM, CREATIS, Villeurbanne, France.}
\address[criugm]{Centre de Recherche de l'Institut de G\'eriatrie de Montr\'eal CRIUGM, Montr\'eal, QC, Canada.}
\address[isi]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA.}

\begin{abstract}
  We investigate six software architectures commonly used to integrate
  workflow engines in science gateways. In \emph{tight integration},
  the workflow engine shares software components with the science
  gateway. In \emph{service invocation}, the engine is isolated and
  invoked through a specific software interface. In \emph{task
    encapsulation}, the engine is wrapped as a computing task executed
  on the infrastructure. In the \emph{pool model}, the engine is
  bundled in an agent that connects to a central pool to fetch and
  execute workflows. In \emph{nested workflows}, the engine is
  integrated as a child process of another engine. In \emph{workflow
    import}, the engine is integrated through workflow language
  conversion. We describe and evaluate these architectures with
  metrics for assessment of integration complexity, robustness,
  extensibility, scalability and support for meta-workflows and
  fine-grained debugging\closednote{Naj}{Please be precise, what are
    they?}{Tristan}{Done, thanks}. Results provide insights for
  science gateway architects and workflow engine developers. Tight
  integration and task encapsulation are the easiest to integrate and
  the most robust. Extensibility is equivalent in most
  architectures. The pool model is the most scalable one and
  meta-workflows are only available in nested workflows and workflow
  import. \closednote{Naj}{I find this abstract to be vague, it can be made
    more concise and poignant}{Tristan}{Added a description of the
    architectures, following @pbellec's suggestion.}
\end{abstract}

@@

\note{sil}{should we then also thank SURF foundation (for the dutchgrid), FP7 (for SCI_BUS, SHIWA and ER-flow), and COMMIT/ and HUvA HPCN for the funding for the amc science gateway activities. My participation has been minimal, and the amc gateways are not discussed in detail (my fault), so maybe it would be an overkill}

>>>>

\author[mcgill]{Reza Adalat}
\author[mcgill]{Natacha Beck}
\author[mcgill]{\\Samir Das}
\author[isi]{Rafael Ferreira da Silva}
\author[mcgill]{Najmeh Khalili-Mahani}
\author[criugm]{Pierre-Olivier Quirion}
\author[mcgill]{Pierre Rioux}
\author[amc]{Si\'lvia D. Olabarriaga}
\author[criugm]{Pierre Bellec}
\author[mcgill]{Alan C. Evans}

\address[mcgill]{McGill Centre for Integrative Neuroscience, Montreal Neurological Institute, McGill University, Canada.}
\address[creatis]{University of Lyon, CNRS, INSERM, CREATIS, Villeurbanne, France.}
\address[criugm]{Centre de Recherche de l'Institut de G\'eriatrie de Montr\'eal CRIUGM, Montr\'eal, QC, Canada.}
\address[isi]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA.}
\address[amc]{Academic Medical Center of the University of Amsterdam, Department of Clinical Epidemiology, Biostatistics and Bioinformatics, Amsterdam, NL.}

\begin{abstract}
	\note{sil}{i find this a dry start. what about an intro sentence?
		Workflow engines are critical for the efficient and transparent
		exploitation of distributed infrastructures in the ecosystem of tools
		and services offered by science gateways.}
  We investigate six software architectures commonly used to integrate
  workflow engines into science gateways. In \emph{tight integration},
  the workflow engine shares software components with the science
  gateway. In \emph{service invocation}, the engine is isolated and
  invoked through a specific software interface. In \emph{task
    encapsulation}, the engine is wrapped as a computing task executed
  on the infrastructure. In the \emph{pool model}, the engine is
  bundled in an agent that connects to a central pool to fetch and
  execute workflows. In \emph{nested workflows}, the engine is
  integrated as a child process of another engine. In \emph{workflow
    import}, the engine is integrated through workflow language
  conversion. We describe and evaluate these architectures with
  metrics for assessment of integration complexity, robustness,
  extensibility, scalability and support for meta-workflows and
  fine-grained debugging\closednote{Naj}{Please be precise, what are
    they?}{Tristan}{Done, thanks}. Results provide insights for
  science gateway architects and workflow engine developers. 
  \note{sil}{what are the insights for wf engine developers?. this is not clear in the discussion and conclusions. i think the focus lies more on the integration and not on one or the other} Tight
  integration and task encapsulation are the easiest to integrate and
  the most robust. Extensibility is equivalent in most
  architectures. The pool model is the most scalable one and
  meta-workflows are only available in nested workflows and workflow
  import. \closednote{Naj}{I find this abstract to be vague, it can be made
    more concise and poignant}{Tristan}{Added a description of the
    architectures, following @pbellec's suggestion.}
\note{sil{what about closing with this sentence from above:
		These results provide insights for
		science gateway architects and workflow engine developers. }
\end{abstract}

@@

\note{sil}{should we then also thank the dutch crowd?
	 We also acknowledge the Dutch national e-Infrastructure with the support of SURF Cooperative,  the Dutch national program COMMIT/ and the High Performance Computing and Networking (HPCN) Fund of the University of Amsterdam for their support to the science gateway actitivies at the AMC. 
	 We are also greatful to the financial support provided by FP7 E-INFRASTRUCTURE program for financial support to SCI_BUS, SHIWA and ER-flow projects.  
	 My participation has been minimal, and the amc gateways are not discussed in detail (my fault), so maybe it would be an overkill}
